Two Decades of Smalltalk VM Development

### 2018-08-17 paper.tex
nuke a spurious "out"

\paragraph{Simulation memory layout.}In the simulator, the heap is stored as a large contiguous byte array. References between objects are indices inside the byte array instead of pointers. All the Slang variables, normally translated to C variables, are simulated as Smalltalk objects. They use specific wrapper classes, such as \emph{CArrayAccessor} over normal Smalltalk classes, to emulate the C behavior (only array accesses are available in C, not high level iterator APIs, etc.). The Slang code is executed as Smalltalk code. The Smalltalk stack zone is represented in \openSmalltalkVM as a doubly linked cycle of stack pages which are maintained by the VM. In the interpreter-only simulator, the stack zone is a Smalltalk Array object. In the full simulator, the stack zone is in the byte array, to allow access to the stack from generated machine code by the processor simulator\footnote{This difference is historical, the interpreter-only version was implemented first in the easiest way possible, but out with the JIT the Smalltalk stack must be accessible to the processor simulator.}.

>>>>

\paragraph{Simulation memory layout.}In the simulator, the heap is stored as a large contiguous byte array. References between objects are indices inside the byte array instead of pointers. All the Slang variables, normally translated to C variables, are simulated as Smalltalk objects. They use specific wrapper classes, such as \emph{CArrayAccessor} over normal Smalltalk classes, to emulate the C behavior (only array accesses are available in C, not high level iterator APIs, etc.). The Slang code is executed as Smalltalk code. The Smalltalk stack zone is represented in \openSmalltalkVM as a doubly linked cycle of stack pages which are maintained by the VM. In the interpreter-only simulator, the stack zone is a Smalltalk Array object. In the full simulator, the stack zone is in the byte array, to allow access to the stack from generated machine code by the processor simulator\footnote{This difference is historical, the interpreter-only version was implemented first in the easiest way possible, out with the JIT the Smalltalk stack must be accessible to the processor simulator.}.
